---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# STICr

<!-- badges: start -->
<!-- badges: end -->

The goal of STICr (pronounced "sticker") is to provide a standardized set of functions for working with data from Stream Temperature, Intermittency, and Conductivity (STIC) loggers, first described in [Chapin et al. (2014)](https://agupubs.onlinelibrary.wiley.com/doi/abs/10.1002/2013WR015158). STICs and other intermittency sensors are becoming more popular, but their raw data output is not in a form that allows for convenient analysis. This package aims to provide a set of functions for tidying the raw data from these loggers, as well as calibrating their conductivity measurements to specific conductivity (`SpC`) and classifying the conductivity data to generate a classified "wet/dry" data set.

## Installation

You can install STICr from CRAN or the development version of STICr from [GitHub](https://github.com/HEAL-KGS/STICr) with:

```{r setup}
# install.packages("STICr")  # if needed: install package from CRAN
# devtools::install_github("HEAL-KGS/STICr") # if needed: install dev version from GitHub
library(STICr)
```

## Example

This is an example workflow that shows the main functionality of the package. A more detailed version is available in the package vignette. 

### Step 1: Load data

```{r load-data}
# read in raw HOBO data and tidy
df_tidy <- tidy_hobo_data(infile = "https://samzipper.com/data/raw_hobo_data.csv", outfile = FALSE)
head(df_tidy)
```

### Step 2: Get and apply calibration

The second function is called `get_calibration` and is demonstrated below. The function intakes a STIC calibration data frame with columns `standard` and `conductivity_uncal`and outputs a fitted model object relating `spc` to the uncalibrated conductivity values measured by the STIC.

```{r calibrate}
# load calibration
lm_calibration <- get_calibration(calibration_standard_data)

# apply calibration
df_calibrated <- apply_calibration(
  stic_data = df_tidy,
  calibration = lm_calibration,
  outside_std_range_flag = T
)
head(df_calibrated)
```

### Step 3: Classify data

```{r classify-data}
# classify data
df_classified <- classify_wetdry(
  stic_data = df_calibrated,
  classify_var = "SpC",
  threshold = 100,
  method = "absolute"
)
head(df_classified)
```

### Step 4: QAQC

```{r qaqc-data}
# apply qaqc function
df_qaqc <-
  qaqc_stic_data(
    stic_data = df_classified,
    spc_neg_correction = T,
    inspect_classification = T,
    anomaly_size = 2,
    window_size = 96,
    concatenate_flags = T
  )
head(df_qaqc)
table(df_qaqc$QAQC)
```

### Step 5: Plot classified data

```{r plot-classified-data, fig.width = 5, fig.height = 3}
# plot SpC through time, colored by wetdry
plot(df_classified$datetime, df_classified$SpC,
  col = as.factor(df_classified$wetdry),
  pch = 16,
  lty = 2,
  xlab = "Datetime",
  ylab = "Specific conductivity"
)
legend("topright", c("dry", "wet"),
  fill = c("black", "red"), cex = 0.75
)
```

### Step 6: Compare to field observations

```{r validate-data}
# create validation data frame
stic_validation <-
   validate_stic_data(
     stic_data = classified_df,
     field_observations = field_obs,
     max_time_diff = 30,
     join_cols = NULL,
     get_SpC = TRUE,
     get_QAQC = FALSE)

# compare the field observations and classified STIC data in table
head(stic_validation)

# calculate percent classification accuracy
sum(stic_validation$wetdry_obs == stic_validation$wetdry_STIC)/length(stic_validation$wetdry_STIC)

# compare SpC as a plot
plot(stic_validation$SpC_obs, stic_validation$SpC_STIC,
     xlab = "Observed SpC", ylab = "STIC SpC")

```
