---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# STICr

<!-- badges: start -->
<!-- badges: end -->

The goal of STICr is to provide a standardized set of functions for working with data from Stream Temperature, Intermittency, and Conductivity (STIC) loggers, first described in [Chapin et al. (2014)](http://onlinelibrary.wiley.com/doi/10.1002/2013WR015158/abstract). STICs and other intermittency sensors are becoming more popular, but their raw data output is not in a form that allows for convenient analysis. This package aims to provide a set of functions for tidying the raw data from these loggers, as well as calibrating their conductivity measurements to specific conductivity (`spc`). Additionally, the `classify_wetdry` function allows the user to define both a method and threshold value to generate a binary "wet/dry" data set.

## Installation

You can install the development version of STICr from [GitHub](https://github.com/HEAL-KGS/STICr) with:

``` r
# install.packages("devtools")
devtools::install_github("HEAL-KGS/STICr")
```

## Example

This is an example workflow that shows the main functionality of the package. This first function is called `tidy_hobo_data` and is demonstrated below. The function intakes a raw CSV file and outputs a tidy data frame with the following column names: `datetime`, `conductivity_uncal`, and `temperature`.

```{r example}
library(STICr)

clean_data <- tidy_hobo_data(infile = "https://raw.githubusercontent.com/HEAL-KGS/STICr/main/data/raw_hobo_data.csv", outfile = FALSE)
head(clean_data)
```

Below are time series plots of the uncalibrated conductivity and temperature generated by `tidy_hobo_data`

```{r plot}
plot(clean_data$datetime, clean_data$conductivity_uncal, type = "l", lty = 1, 
     xlab = "datetime", ylab = "uncalibrated conductivity")

plot(clean_data$datetime, clean_data$temperature, type = "l", lty = 1, 
     xlab = "datetime", ylab = "temperature")
```

The second function is called `get_calibration` and is demonstrated below. The function intakes a STIC calibration data frame with columns `standard` and `conductivity_uncal`and outputs a fitted model object relating `spc` to the uncalibrated conductivity values measured by the STIC.

```{r example2}
# Example calibration data
head(calibration_standard_data)

# Creating model object using a linear interpolation method. User can also chose the "exponential" method
lm_calibration <- get_calibration(calibration_standard_data, method = "linear")

# Result is a fitted model object
summary(lm_calibration)
```

The third function is called `apply_calibration` and is demonstrated below. The function intakes a tidy STIC data frame (as generated by` tidy_hobo_data`), as well as the fitted model object generated by `get_calibration`. It outputs the same STIC data frame as input, except with a new column called `spc`. This will be in the same units as the data used to develop the model calibration.

```{r example3}
# applying the lm_calibration object from the previous example to the tidy data frame to generate the calibrated spc column
calibrated_df <- apply_calibration(clean_data, lm_calibration)
head(calibrated_df)
```

Below is a time series plot of the calibrated spc values

```{r plot2}
plot(calibrated_df$datetime, calibrated_df$spc, type = "l", lty = 1, 
     xlab = "datetime", ylab = "specific conductivity")
```

The final function is called `classify_wetdry` and is demonstrated below. This function intakes a data frame with STIC data (such as that produced by `apply_calibration` or `tidy_hobo_data.` The user is also asked to specify a method for classification, as well as a threshold associated with that method. Currently, methods include either an absolute numerical threshold or a percentage of the highest value of the classification variable. It outputs the same STIC data frame as input, except with a new binary column called `wetdry`.

```{r example4}
# Creating classified data frame using "spc" as the classification variable, and an absolute threshold as the method for classification. The spc threshold chosen in this case is 200. 
classified_df <- classify_wetdry(calibrated_df, classify_var = "spc", method = "absolute",
                                 threshold = 200)
head(classified_df)
```

Below is a time series plot of the calibrated `spc` values colored by wet versus dry periods.

```{r plot3}
classified_df$wetdry <- as.factor(classified_df$wetdry)
plot(classified_df$datetime, classified_df$spc, col = classified_df$wetdry, lty = 1, type = "l",  
     xlab = "datetime", ylab = "specific conductivity")
```
