---
title: "STIC data processing and report"
author: "Christopher Wheeler"
date: "4/6/2022"
output:  rmarkdown::github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## STIC Processing Workflow

```{r packages, echo = T, results = 'hide', error=FALSE, warning= FALSE, message = FALSE}

library(tidyverse)
library(cowplot)
library(dygraphs)
library(patchwork)
library(sf)

```
I have developed some generalized functions for tidying and preliminary processing of STIC data that I intend to make into a package. Code for the functions is below:

The first function loads raw HOBO STIC files and cleans up columns and headers
to produce a tidy data frame and/or CSV output.

infile = filename (including path if needed) for the raw HOBO data on your computer.

outfile = filename (including path if needed) to save the tidied. Defaults to FALSE, in which case tidied data will not be saved.


```{r STICr Function 1}

source("tidy_hobo_data.R")

```
This next function (get_calibration.R) is a function to fit standard and measured calibration data to model a object (lm and exponential to start)

inputs: data frame with columns "standard" and "conductivity_uncal" and method = "linear" or "exponential"

output: fitted model object

input needs to be data frame. column names: conductivity_uncal, standard
```{r STICr Function 2}

source("get_calibration.R")

```
The next function is called apply_calibration.R

inputs: calibration (a model object) and a data frame with columns labeled "datetime", "temperature" "conductivity_uncal"

output: same data frame as input, except with a new column called "spc"
```{r STICr Function 3}
source("apply_calibration.R")
```
Last function is classify_wetdry.R

inputs: data frame with columns "datetime", "temperature", "conductivity", and "specific_conductivity"

output: same data frame as input, but with a new column called wetdry

```{r STICr Function 4}
source("classify_wetdry.R")
```

## Examples of the use of STICr functions

Here is how they work:

First we are bringing in and tidying raw HOBO data by giving the function a filepath. In this example it is Youngmeyer STIC ENM312 from July to February

```{r Testing_1}

stic_data <- tidy_hobo_data("C:/Users/cwhee/Desktop/R_Directory/AIMS_hydro_qaqc/youngmeyer_stics_round_1/21038945_ENM312.csv")

head(stic_data)

```
Next we are bringing in a set of calibration points and creating a fitted model object. In this case, the model type we chose is linear 

```{r Testing_2}

cal_points <- read_csv("cal_points.csv")

cal_points <- cal_points %>% 
  rename(standard = std_val) %>% 
  rename(conductivity_uncal = measured_val)

calibration_data <- cal_points

calibration <- get_calibration(calibration_data, method = "linear")

summary(calibration)
```
Now Step 3: Apply calibration. The function take the current data frame and model object, then adds a new "real" spc column to data frame

```{r Testing_3}

stic_data_with_spc <- apply_calibration(input = stic_data, calibration = calibration)

head(stic_data_with_spc)
```
Final step: using the classify_wetdry function to create another column with the binary, classified data. Default threshold is 1000 for conductivity_uncal

```{r Testing_4}

classified <- classify_wetdry(stic_data_with_spc, threshold = 50)

head(classified)

```

Last thing: we need to change "negative" calibrated spc values to 0

```{r fix_spc}

classified$spc <- ifelse(classified$spc < 0, 0, classified$spc)

```

## What does this data (for one stic) look like

Here is the uncalibrated conductivity time series:

```{r graphs, echo = TRUE}

ggplot(classified, aes(x = datetime, y = spc, color = wetdry, group = 1)) +
  geom_path(size = 1) + 
  theme_cowplot()

```

Here is a temperature time series

```{r graphs_2, echo = TRUE}

ggplot(classified, aes(x = datetime, y = temperature)) +
  geom_path() + 
  geom_smooth(color = "red") +
  theme_cowplot()

```

Here is some code showing how to bring all the STIC locations in and calculate metrics:

```{r stics full packages, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE}

library(fs)
library(Rcpp)
library(sp)
library(raster)
library(rgdal)
library(rasterVis)
library(sf)

```

```{r stics full, echo=TRUE,warning=FALSE,message=FALSE,error=FALSE}

data_dir <- "processed_files_youngmeyer"

fs::dir_ls(data_dir)

length(data_dir)

stic_files <- fs::dir_ls(data_dir, regexp = "\\.csv$")

### using the map_dfr function from tidyverse to loop in individual csvs and bind rows
stic_files <- stic_files %>% 
  map_dfr(read_csv)

### Creating binary wet/dry column based of relative cond value of 1000
stic_files <- stic_files %>% 
  dplyr::mutate(wetdry = if_else(conductivity >= 1000, "wet", "dry" ))

### using lubridate to convert datetime to tidy format 
stic_files <- stic_files %>% 
  mutate(datetime = lubridate::mdy_hms(datetime))

### Cutting off dates to begin at 07/18/2021
stic_files <- stic_files %>% 
  filter(datetime >= "2021-07-18")

### using lubridate to create column with just date rather than datetime 
stic_files <- stic_files %>% 
  mutate(date = lubridate::date(datetime))

### using lubridate to separate out the individual days 
stic_files <- stic_files %>% 
  mutate(day = lubridate::yday(date))


### making a new dataframe with date of first drying for all individual logger
drying <- stic_files %>% 
  subset(wetdry == "dry") %>%
  group_by(logger) %>% 
  summarise(first_dry = min(day)) 

### Now calculating the percentage of wet STICs at each time step
### Calculate wet network proportion

stic_wet_prop <- stic_files %>% 
  group_by(datetime) %>% 
  summarise(n_wet = sum(wetdry == "wet"), n_sensors = n() ) %>% 
  mutate(percent = n_wet/n_sensors)

ggplot(stic_wet_prop, aes(x = datetime, y = percent)) + 
  geom_line(color = "steelblue", size = 1) + 
  ylab("wet network proportion") + 
  theme_cowplot()
  

### Calculate flow duration for each sensor
### i,e, percentage of wet observations out of total obs for each logger

stic_duration <- stic_files %>% 
  group_by(logger) %>% 
  summarise(n_wet = sum(wetdry == "wet", na.rm = TRUE), count = n()) %>% 
  mutate(duration = n_wet/count)

ggplot(stic_duration, aes(x = duration)) + 
  geom_histogram(fill = "steelblue") + 
  theme_cowplot() + 
  xlab("flow duration for logger") + 
  ylab("count of loggers")

ggplot(stic_duration, aes(x = duration)) + 
  geom_density(color = "steelblue", size = 1) + 
  theme_cowplot()



```
Now we are pairing with a coordinate data frame to create some maps. First we are just going to map the location of ENM312

```{r maps full, echo=TRUE}

library(tidyverse)
library(tidyr)

youngmeyer_coords <- read_csv("youngmeyer_coords.csv", 
                              col_types = cols(logger = col_number()))

stic_duration$logger <- as.numeric(stic_duration$logger)

duration_coords <- left_join(youngmeyer_coords, stic_duration, by = "logger")

### Make Duration Map


youngmeyer_duration_locs <- st_as_sf(duration_coords,
                                     coords = c("Longitude", "Latitude"), 
                                     crs = 4326)


youngmeyer_cal <- read_csv("youngmeyer_cal.csv", 
                           col_types = cols(logger = col_number()))


young_spc_stic <- left_join(stic_files, youngmeyer_cal, by = "logger")

young_spc_stic <- drop_na(young_spc_stic)

young_spc_stic <- young_spc_stic %>% 
  mutate(spc = conductivity/slope)


young_spc_wet_stic <- young_spc_stic %>% 
  filter(wetdry == "wet")

# generate average cond value
young_ave_cond <- young_spc_wet_stic %>% 
  group_by(logger) %>% 
  summarise(average_spc = mean(spc)) 

# map the mean spc

youngmeyer_coords <- read_csv("youngmeyer_coords.csv", 
                              col_types = cols(logger = col_number()))


spc_coords <- left_join(youngmeyer_coords, young_ave_cond, by = "logger")

### Make SPC Map

youngmeyer_spc_locs <- st_as_sf(spc_coords,
                                coords = c("Longitude", "Latitude"), 
                                crs = 4326)

# Put in stream network shapefile


youngmeyer_streams <- st_read(
  "youngmeyer_shape/Youngmeyer_StreamNetwork.shp")

# Make single map

enm312_loc <- youngmeyer_duration_locs %>% 
  filter(Site == "ENM312")


ggplot() + 
  geom_sf(data = youngmeyer_streams, size = 1) +
  geom_sf(data = enm312_loc, size = 4) +
  scale_color_viridis_c() +
  theme_classic() +
  xlab("Longitude") + 
  ylab("Latitude")



# Make New Maps with Stream Lines

# ave spc
ggplot() + 
  geom_sf(data = youngmeyer_streams, size = 1) +
  geom_sf(data = youngmeyer_spc_locs, aes(color = average_spc), size = 4) +
  scale_color_viridis_c() +
  theme_classic() +
  xlab("Longitude") + 
  ylab("Latitude")



ggplot() + 
  geom_sf(data = youngmeyer_streams, size = 1) +
  geom_sf(data = youngmeyer_duration_locs, aes(color = duration), size = 4) +
  scale_color_viridis_c() +
  theme_classic() +
  xlab("Longitude") + 
  ylab("Latitude")

```



